Word learning at the semantics pragmatics interface

### 2020-04-16 paper/supplementary_material.Rmd
edits to model parameters sections, add regression math

Our model describes a listener reasoning about the referent referred to by a speaker's utterance. This reasoning is contextualized by the prior probability of each referent $P(r \mid \rho_i)$ . This prior probability is thought to be a function of the common ground $\rho$ shared between speaker and listener in that interacting around the objects changes the probability that they will be referred to later. We assume that the degree to which interactions around objects are integrated into the common ground and thus change the prior probability of those objects depends on the child's age $i$.

@@

The informativity of each utterance is given by imagining which referent a literal listener, who interprets words according to their lexicon $\mathcal{L}$, would infer upon hearing the utterance. Thus, this reasoning depends on what kind of semantic knowledge (word--object mappings) the speaker thinks the literal listener knows. We parameterize the  listener's knowledge of a word's semantics in terms of a semantic knowledge parameter $\theta$, which varies between 0 and 1. $\theta = 0$ correspondes to the state of knowledge for a completely novel word and results in a semantic interpretation function that chooses randomly between the objects in the scene. Each of the novel objects are assumed to have semantic knowledge of 0. For $\theta \in (0, 1)$, the semantic interpretation function will pick out the familiar referent with probability $\theta + \frac{1 - \theta}{2} = \frac{1 + \theta}{2}$; that is, with probability $\theta$, the listener knows the correct meaning of the word (and picks out the correct referent 100\% of the time); with  probability $1 - \theta$, the listener does not know the meaning of the word and must guess, picking out the correct referent 50\% of the time. For familiar objects, semantic knowledge is a function of the degree-of-acquisition of the associated word, which in turn depends upon the kind of object $j$ (its expected acquisition trajectory) as well as on the child's age $i$.

@@

The model description above points to three potential loci of developmental change: semantic knowledge, expectations about speaker informativeness and sensitivity to common ground. Each of theses components is represented by a parameter that plays a particular functional role in the model. We capture developmental change by making these parameters a function of age and therefore estimating a developmental trajectory (intercept and slope) for each parameter.

@@

Semantic knowledge captures the degree of certainty with which the naive listener is assumed to know the label for the familiar object. As a consequence, semantic knowledge differs between familiar objects. For objects whose labels are generally acquired earlier (e.g. carrot) semantic knowledge should generally be high whereas for others (e.g. pawn) semantic knowledge should generally be lower. However, semantic knowledge also varies with age such that older children are more likely to know the labels for more of the familiar objects compared to younger children. As a consequence, each familiar object has a unique developmental trajectory with respect to semantic knowledge. Technically, the object-specific parameters ($\theta_{ij}$) are estimated in the form of a hierarchical regression (mixed-effects) model; that is, each object's trajectory is estimated as a deviation from an overall trajectory of vocabulary development. This overall (vocabulary) trajectory represents the development of semantic knowledge that is independent of a particular familiar object--label pairing (see Figure \@ref(fig:bda)).

@@

A second locus of developmental change is a listener's expectations about speaker informativeness. In the context of the model, speaker informativeness corresponds to the degree with which the listener expects the speaker to choose the most informative of the available utterances. We assume that children at different ages could have different expectations about how rational or informative speakers are [see e.g. @bohn_tessler_merrick_frank_2019; @frank2014inferring; @yoon2019role].

@@

Sensitivity to common ground refers to the probability that an object is taken to be the referent of the utterance before actually hearing the utterance. Thus, it captures the salience of an object due to its role in the social interaction that precedes the utterance. We expect children at different ages to respond differently to the common ground manipulation, resulting in an age specific prior distribution over objects [akhtar1996role; @diesendruck2004two].

@@

We used the following prior distributions for model parameters. Intercept and slope for sensitivity to common ground: $unif(-2,2)$. For the speaker informativeness we used $unif(-3,3)$ for the intercept $unif(-0,4)$ for the slope. We restricted the slope to be positive because negative values for speaker informativeness are conceptually implausible. For semantic knowledge we used $unif(-3,3)$ for the intercept and $unif(0,2)$ for the slope, because it is implausible to assume that semantic knowledge decreases with age. For the parameter capturing the variability the object specific trajectories around these overall parameters we used $unif(0,2)$ for the intercept and $unif(0,1)$ for the slope. Some choices regarding these prior distributions were made to ease model convergence. However, please note that all models considered used the same prior distributions.

>>>>

Our model describes a listener reasoning about the referent referred to by a speaker's utterance. This reasoning is contextualized by the prior probability of each referent $P(r \mid \rho_i)$. This prior probability is a function of the common ground $\rho$ shared between speaker and listener in that interacting around the objects changes the probability that they will be referred to later. We assume that the degree to which interactions around objects are integrated into the common ground (and thus change the prior probability of those objects) depends on the child's age $i$.

@@

The informativity of each utterance is given by imagining which referent a literal listener, who interprets words according to their lexicon $\mathcal{L}$, would infer upon hearing the utterance. Thus, this reasoning depends on what kind of semantic knowledge (word--object mappings) the speaker thinks the literal listener knows. We parameterize the  listener's knowledge of a word's semantics in terms of a semantic knowledge parameter $\theta$, which varies between 0 and 1. $\theta = 0$ correspondes to the state of knowledge for a completely novel word and results in a semantic interpretation function that chooses randomly between the objects in the scene. Each of the novel words are assumed to have semantic knowledge of 0. For $\theta \in (0, 1)$, the semantic interpretation function will select the familiar referent with probability $\theta + (1 - \theta)\frac{1}{2} = \frac{1 + \theta}{2}$; that is, with probability $\theta$, the listener knows the correct meaning of the word (and picks out the correct referent 100\% of the time); with probability $1 - \theta$, the listener does not know the meaning of the word and must guess, picking out the correct referent 50\% of the time.^[
  One could also define $\theta$ to directly denote the probability of selecting the familiar~vs.~unfamiliar referent. The model behavior would be the same except for a translation of this parameter. We choose the parametrization described because we find it more intuitive for $\theta$ to represent the degree of semantic knowledge, where 0 semantic knowledge represents a state of complete uncertainty.
] For familiar objects, semantic knowledge is a function of the degree-of-acquisition of the associated word, which in turn depends upon the word $j$ (its expected acquisition trajectory) as well as on the child's age $i$.

@@

The model description above points to three potential loci of developmental change: semantic knowledge, expectations about speaker informativeness and sensitivity to common ground. Each of theses components is represented by a parameter that plays a particular functional role in the model. We capture developmental change by making these parameters a function of age $i$ for which we estimate a developmental trajectory (intercept and slope in a (logistic-) linear model) for each parameter.

@@

Semantic knowledge captures the degree of certainty with which the naive listener is assumed to know the label for the familiar object. As a consequence, semantic knowledge differs among familiar words. For objects whose labels are generally acquired earlier (e.g., "carrot") semantic knowledge should generally be high whereas for others (e.g., "pawn") semantic knowledge should generally be lower. However, semantic knowledge also varies with age such that older children are more likely to know the labels for more of the familiar objects compared to younger children. As a consequence, each familiar word has a unique developmental semantic knowledge trajectory.

Technically, the item-specific parameters ($\theta_{ij}$) are estimated in the form of a hierarchical regression (mixed-effects) model: $\theta_{ij} = \text{logistic}(\beta^{\theta}_{0, j} + i \cdot \beta^{\theta}_{1, j})$; each word's lexical development trajectory (the intercept $\beta^{\theta}_{0, j}$  and slope $\beta^{\theta}_{1, j}$ of the regression line for each object) is estimated as a deviation from an overall trajectory of vocabulary development. The intercept and slope for each item are sampled from Gaussian distributions with means $\mu^\theta_0, \mu^\theta_1$ and variances $\sigma^\theta_0, \sigma^\theta_1$: $\beta^{\theta}_{0, j} \sim \mathcal{N}(\mu^\theta_0, \sigma^\theta_0)$ and $\beta^{\theta}_{1, j} \sim \mathcal{N}(\mu^\theta_1, \sigma^\theta_1)$.
$\mu^\theta_0$ and $\mu^\theta_1$ represent the overall vocabulary development independent of particular familiar object--label pairings, and $\sigma^\theta_0$ and $\sigma^\theta_1$ represent the overall variability of intercepts and of slopes between items (see Figure \@ref(fig:bda)). For all regression components, intercepts correspond to the knowledge or sensitivity for our youngest age bin: 2-year-olds. For example, $\mu^\theta_0$ represents the average semantic knowledge of our 2-year-olds and $\mu^\theta_1$ represents the growth in semantic knowledge over developmental time. $\sigma^\theta_0$ represents the variability in the semantic knowledge of our items (for the 2-year-olds) and $\sigma^\theta_1$ represents the variability in growth of semantic knowledge across our different items.

@@

A second locus of developmental change is a listener's expectations about speaker informativeness $\alpha$. In the context of the model, speaker informativeness corresponds to the degree with which the listener expects the speaker to choose the most informative of the available utterances. We assume that children at different ages could have different expectations about how rational or informative speakers are [see e.g. @bohn_tessler_merrick_frank_2019; @frank2014inferring; @yoon2019role], which we model with a linear function: $\alpha_i = \beta^\alpha_0 + i \cdot \beta^\alpha_1$.

@@

Sensitivity to common ground $\rho$ refers to the probability that an object is taken to be the referent of the utterance before actually hearing the utterance. Thus, it captures the salience of an object due to its role in the social interaction that precedes the utterance. We expect children at different ages to respond differently to the common ground manipulation, resulting in an age specific prior distribution over objects [akhtar1996role; @diesendruck2004two], which we model with a logistic-linear function: $\rho_i = \text{logistic}(\beta^\rho_0 + i \cdot \beta^\rho_1)$.

@@

We used the following prior distributions for model parameters. Intercept and slope for sensitivity to common ground: $\beta^\rho_0 ,\beta^\rho_1 \sim \text{Uniform}(-2,2)$. Speaker informativeness:  $\beta^\alpha_0 \sim \text{Uniform}(-3,3)$ for the intercept and $\beta^\alpha_1 \sim \text{Uniform}(-0,4)$ for the slope. We restricted the slope to be positive because negative values for speaker informativeness are conceptually implausible: We do not expect sensitivity to speaker informativeness to decrease across our age range. For the global semantic knowledge (vocabulary) parameters, we used $\mu^\theta_0 \sim \text{Uniform}(-3,3)$ for the intercept and $\mu^\theta_1 \sim \text{Uniform}(0,2)$ for the slope, because it is implausible to assume that semantic knowledge decreases with age. For the parameters capturing the variability the object specific trajectories around these overall parameters we used $\sigma^\theta_0 \sim \text{Uniform}(0,2)$ for the intercept and $\sigma^\theta_1 \sim \text{Uniform}(0,1)$ for the slope. Some choices regarding these prior distributions were made to ease model convergence. However, please note that all models considered used the same prior distributions.
